Learning Auditory Similarities Between Instrumental Playing Techniques

### 2019-09-24 paper/paper.tex
clarify intro

Despite their widespread use, linguistic references to musical timbre suffer from limited adequacy.
On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them.
On the other hand, the sole mention of playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium.
For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in Italian.

The prospect of retrieving timbre perception across playing techniques necessarily exceeds the symbolic domain.
Instead, it involves a cognitive process which arises from the subjective experience of listening to music.
From the standpoint of audio signal processing, the task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
Rather than to merely discriminate instruments as mutually exclusive categories, this distance function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal.

>>>>

Despite their widespread use, linguistic references to musical timbre suffer from limited adequacy.
On the one hand, adjectives such as \emph{bright} or \emph{rough} are prone to misunderstanding, as they do not prescribe any musical gesture that is capable of achieving them.
On the other hand, the sole mention of playing technique does not determine any intended effect; and therefore, fails to generalize to the entire instrumentarium.
For instance, although the term \emph{breathy} alludes to a playing technique that is peculiar to wind instruments, a cellist may accomplish a seemingly breathy timbre by bowing near the fingerboard, i.e., \emph{sul tasto} in the classical terminology.

The prospect of modeling timbre perception necessarily exceeds the symbolic domain.
Instead, it involves a cognitive process which arises from the subjective experience of listening.
From an audio signal processing standpoint, the task of simulating this cognitive process amounts to the design of a multidimensional feature space wherein some distance function evaluates arbitrary pairs of stimuli.
Rather than to merely discriminate instruments as mutually exclusive categories, this distance function must reflect a consensus of quantitative judgments of acoustic dissimilarity, all other parameters --- \ie{} duration, pitch, and intensity --- being equal.
